Glyph-maps for Visually Exploring Temporal Patterns in Climate Data and Models

### 2012-05-14 revision-final/paper.tex
fixed typos around pages 8,9

Another common approach for displaying spatiotemporal data is to calculate a statistical summary for each location and display these values as single heatmap. For example, to study long-term trends, the slope of a linear model can be calculated at each location and color tiles used to display the slope value over space. This approach is less than exploratory, in the sense that it requires a pre-conceived definition of the relationship of interest, and clean data so that the summary statistic adequately summarizes the pattern.

@@

A possible draw-back of the glyph-map, particularly if linear trend is displayed as an icon, is it may suffer from the Z\"ollner Illusion \citep{Zollner}, which makes straight lines look crooked (Section \ref{sec:large-data} has an example).

@@

Each glyph is small, so there is not enough space for a full set of axes on each, and it is not desirable beause it would be repetitive, redundant information. Instead, minimal reference lines and boxes are incorporated. These need to be minimally perceptible, post-attentive \citep{healey} and de-emphasized, so as not to detract from the data. The reference grid is a box, framing each icon, which represents the spatial grid. The reference line is a horizontal line at mid-range. Both help to read differences in slope and intercept. Slope is read by comparing the position of left and right endpoints on the frame, or by reading the angle between the data line and the reference line. Intercept is read from ``average'' position of the line in the box: is it near the top, or near the bottom? Figure~\ref{fig:ref-basic} demonstrates both guides. Reference elements are drawn in white, so they are minimally perceptible relative to the black data elements. Average temperatures were calculated for each month, and these were passed to the glyph calculations to produce the glyph-maps.   For star glyphs the reference frames are also useful and the equivalent of the reference line is a circle (see Figure~\ref{fig:templates}).

>>>>

Another common approach for displaying spatiotemporal data is to calculate a statistical summary for each location and display these values as asingle heatmap. For example, to study long-term trends, the slope of a linear model can be calculated at each location and color tiles used to display the slope value over space. This approach is less than exploratory, in the sense that it requires a pre-conceived definition of the relationship of interest, and clean data so that the summary statistic adequately summarizes the pattern.

@@

A possible drawback of the glyph-map, particularly if linear trend is displayed as an icon, is it that may suffer from the Z\"ollner Illusion \citep{Zollner}, which makes straight lines look crooked (Section \ref{sec:large-data} has an example).

@@

Each glyph is small, so there is not enough space for a full set of axes on each, and it is not desirable because it would be repetitive, redundant information. Instead, minimal reference lines and boxes are incorporated. These need to be minimally perceptible, post-attentive \citep{healey} and de-emphasized, so as not to detract from the data. The reference grid is a box, framing each icon, which represents the spatial grid. The reference line is a horizontal line at mid-range. Both help to read differences in slope and intercept. Slope is read by comparing the position of left and right endpoints on the frame, or by reading the angle between the data line and the reference line. Intercept is read from ``average'' position of the line in the box: is it near the top, or near the bottom? Figure~\ref{fig:ref-basic} demonstrates both guides. Reference elements are drawn in white, so they are minimally perceptible relative to the black data elements. Average temperatures were calculated for each month, and these were passed to the glyph calculations to produce the glyph-maps.   For star glyphs the reference frames are also useful and the equivalent of the reference line is a circle (see Figure~\ref{fig:templates}).
